---
format: acm-pdf

# use keep-tex to cause quarto to generate a .tex file
# which you can eventually use with TAPS
keep-tex: true

params:
  eval_models: false
  
knitr:
  opts_chunk: 
    cache_comments: false
    crop: true
    
execute: 
  echo: false
  warning: false
  message: false
  include: false

bibliography: beliefs-atypical.bib

title: Changing Beliefs About Correlations in Atypical Scatterplots

# if short-title is defined, then it's used
short-title: 

author:
  - name: Gabriel Strain
    email: gabriel.strain@manchester.ac.uk
    orcid: 0000-0002-4769-9221
    affiliation:
      name: Department of Computer Science, Faculty of Science and Engineering, University of Manchester
      address: Oxford Road
      city: Manchester
      country: United Kingdom
      postal-code: M13 9PL
  - name: Andrew J. Stewart
    email: andrew.j.stewart@manchester.ac.uk
    affiliation:
      name: Department of Computer Science, Faculty of Science and Engineering, University of Manchester
      address: Oxford Road
      city: Manchester
      country: United Kingdom
      postal-code: M13 9PL
  - name: Paul Warren
    email: paul.warren@manchester.ac.uk
    affiliation:
      name: Division of Psychology, Communication and Human Neuroscience, School of Health Sciences, Faculty of Biology, Medicine, and Health, University of Manchester
      address: Oxford Road
      city: Manchester
      country: United Kingdom
      postal-code: M13 9PL
  - name: Charlotte Rutherford
    email: charlotte.rutherford-2@postgrad.manchester.ac.uk
    affiliation:
      name: Division of Psychology Communication and Human Neuroscience, School of Health Sciences, Faculty of Biology, Medicine, and Health, University of Manchester
      address: Oxford Road
      city: Manchester
      country: United Kingdom
      postal-code: M13 9PL
  - name: Caroline Jay
    affiliation:
      name: Department of Computer Science, Faculty of Science and Engineering, University of Manchester
      address: Oxford Road
      city: Manchester
      country: United Kingdom
      postal-code: M13 9PL

# acm-specific metadata
acm-metadata:
  # comment this out to make submission anonymous
  # anonymous: true

  # comment this out to build a draft version
  final: false

  # comment this out to specify detailed document options
  # acmart-options: sigconf, review  

  # acm preamble information
  copyright-year: 2018
  acm-year: 2018
  copyright: acmcopyright
  doi: XXXXXXX.XXXXXXX
  conference-acronym: "Conference acronym 'XX"
  conference-name: |
    Make sure to enter the correct
    conference title from your rights confirmation emai
  conference-date: June 03--05, 2018
  conference-location: Woodstock, NY
  price: "15.00"
  isbn: 978-1-4503-XXXX-X/18/06

  # if present, replaces the list of authors in the page header.
  shortauthors: Strain et al.

  # The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
  # Please copy and paste the code instead of the example below.
  ccs: |
    \begin{CCSXML}
    <ccs2012>
     <concept>
      <concept_id>10010520.10010553.10010562</concept_id>
      <concept_desc>Computer systems organization~Embedded systems</concept_desc>
      <concept_significance>500</concept_significance>
     </concept>
     <concept>
      <concept_id>10010520.10010575.10010755</concept_id>
      <concept_desc>Computer systems organization~Redundancy</concept_desc>
      <concept_significance>300</concept_significance>
     </concept>
     <concept>
      <concept_id>10010520.10010553.10010554</concept_id>
      <concept_desc>Computer systems organization~Robotics</concept_desc>
      <concept_significance>100</concept_significance>
     </concept>
     <concept>
      <concept_id>10003033.10003083.10003095</concept_id>
      <concept_desc>Networks~Network reliability</concept_desc>
      <concept_significance>100</concept_significance>
     </concept>
    </ccs2012>
    \end{CCSXML}
    
    \ccsdesc[500]{Computer systems organization~Embedded systems}
    \ccsdesc[300]{Computer systems organization~Redundancy}
    \ccsdesc{Computer systems organization~Robotics}
    \ccsdesc[100]{Networks~Network reliability}

  keywords:
    - belief change
    - correlation perception
    - scatterplot
    - crowdsourced
  
abstract: |
  abstract goes here

---

```{r}
#| label: setup

set.seed(1234) # random seed for number generation

library(tidyverse)
library(MASS)
library(emmeans)
library(scales)
library(buildmer)
library(lme4)
library(kableExtra)
library(papaja)
library(qwraps2)
library(lmerTest)
library(ggdist)
library(ggpubr)
library(conflicted)
library(ggtext)
library(r2glmm)
library(grid)
library(DescTools)
library(Matrix)

# fix conflicts now using the conflicted package

conflicts_prefer(dplyr::select(), dplyr::filter(), lme4::lmer(), DescTools::AUC())

```

```{r}
#| label: lazyload-cache

if (!params$eval_models){ lazyload_cache_dir("beliefs_attitudes_atypical_scatterplots_cache/pdf") }
```

```{r}
#| label: load-data

# load in data csvs for pre-test and main experiment

pre_test_anon <- read_csv("data/pre_test_response_final.csv")
```

```{r}
#| label: wrangle-data

## NB: With the exception of anonymization, data are provided as-is from 
## pavlovia (survey tool). Wrangling function *must* be run first to make
## the data set usable

wrangle_pre_test <- function(anon_file) {
  
# extract demographic information
  
demographics <- anon_file %>%
  filter(!is.na(gender_slider.response)) %>%
  select(matches(c("participant",
                   "age_textbox.text",
                   "gender_slider.response")))

# select relevant columns
# select only experimental items
# change data types where appropriate
# output this file with suffix 'tidy'

anon_file %>%
  select(c("participant",
           "avg_corr",
           "avg_emot",
           "slider_emotion.response",
           "slider_belief.response",
           "statement",
           "item_no",
           "label",
           "session"
           )) %>%
  filter(item_no < 26) %>%
  full_join(demographics, by = "participant") %>%
  select(-c("__participant")) %>%
  assign(paste0(unique(anon_file$expName), "_tidy"),
           value = ., envir = .GlobalEnv)
}

# use wrangling function on anonymised data file

wrangle_pre_test(pre_test_anon)

# check for missing age and gender id values

sum(is.na(beliefs_scatterplots_pretest_tidy$gender_slider.response))

sum(is.na(beliefs_scatterplots_pretest_tidy$age_textbox.text))

# missing age and gender values for a single participant
# print session with missing values

beliefs_scatterplots_pretest_tidy %>%
  filter(is.na(age_textbox.text)) %>%
  distinct(session)

# session id was used to look up age and gender identity and manually add
# this information to the final results csv using the prolific-supplied
# demographic information for the pretest
# this is not included as would expose personally identifiable prolific IDs

```

```{r}
#| label: author-pre-test-ratings

# NB: this code is reproduced here so that test values are accessible
# It can also be found at item_preparation/pre_test_stim.R

# Create dataframes for each rater

author_A_statements <- read_csv("item_preparation/statements_A.csv") %>%
  rename(Topic_Emotionality_A = Topic_Emotionality,
         Strength_of_Correlation_A = Strength_of_Correlation) %>%
  select(-Notes)
  

author_B_statements <- read_csv("item_preparation/statements_B.csv") %>%
  rename(Topic_Emotionality_B = Topic_Emotionality,
         Strength_of_Correlation_B = Strength_of_Correlation) %>%
  select(-Notes)

# Bind dataframes together

all_ratings <- left_join(author_A_statements, author_B_statements, by = c("Number", "Statements"))

## IRR Calculations

kappa2(matrix(c(all_ratings$Topic_Emotionality_A,
                all_ratings$Topic_Emotionality_B), ncol = 2), "squared")    

kappa2(matrix(c(all_ratings$Strength_of_Correlation_A,
                all_ratings$Strength_of_Correlation_B), ncol = 2), "squared")    

```


# Introduction

# Related Work

# Pre-Study: Investigating Beliefs About Relatedness Statements

## Introduction

Due to previous evidence suggesting effects of prior belief strength and topic
emotionality on the propensity for belief change, we first aim to build a picture
of people's thoughts and feelings along these dimensions in our population of interest.
With the intention of testing the potential for changes in beliefs about correlations
displayed in scatterplots depicting weak and strong correlations, and those whose
topics were both strong and neutral in emotional valence, we began by using
ChatGPT4 [@chatgpt] to generate 100 correlation statements using the following prompt:

\begin{center}

    ``Generate 100 statements that describe the correlation between two variables, such as :

     "X is associated with a higher level of Y" or

     "As X increases, Y increases".

    Try to match all the statements on emotionality.``
\end{center}

The full list of these statements can be found in the supplementary materials. Note
that we cite our use of ChatGPT according to the AI Code of Conduct developed by
Iliada Eleftheriou and Ajmal Mubarik and the University of Manchester [@iliada_2023].
Two authors rated each statement on topic emotionality and strength
of correlation using Likert scales from 1 to 7. Topic emotionality
had a midpoint at 4, whereas strength of correlation went from 1 (Not Related At All)
to 7 (Strongly Related).


## Method

### Participants

### Design

### Procedure

## Results

## Discussion

# Main Study: Potential for Belief Change Using Atypical Scatterplots

## Introduction

## Method

### Participants

### Design

### Procedure

## Results

## Discussion

# General Discussion

# Limitations

# Future Work

# Conclusion

# References {-}




















